### 2020-07-21 Model_Build.Rmd
Fixed proportions on TVMarketSize

---
title: "Model Building"
author: "Jake Esprabens"
date: "7/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fable)
library(lubridate)
library(tsibble)
library(feasts)
library(zoo)
```

## Loading Data
```{r}
gt <- read.csv("MLB_Google_Trends.csv")
ar <- read.csv("MLB_Attendance_and_Revenue.csv")
po <- read.csv("MLB_Preseason_Odds.csv")
tp <- read.csv("MLB_Team_Performance.csv")
tms <- read.csv("MLB_TV_Market_Size.csv")
```

## Cleaning Google Trends Data
```{r}
# cleaning google trends data by replacing <1 with 0 and making hits numeric
# also converting to tsibble
clean_gt <- gt %>%
  mutate(date = as.character(date),
         year = as.character(year(date)),
         hits = str_replace(hits, "<1","0"),
         hits = as.numeric(hits)) %>%
  rename(team = teams,
         googlehits = hits) %>%
  select(-X) %>%
  arrange(team)
```

## Cleaning Preseason Odds Data
```{r}
clean_po <- po %>%
  pivot_longer(cols = 2:18, names_to = "year") %>%
  mutate(year = substring(year, 2)) %>%
  rename(team = Team,
         odds = value)
```

## Cleaning TV Market Size
```{r}
clean_tms <- tms %>%
  mutate_all(na_if,"") %>%
  mutate(TVMarketSize = as.numeric(gsub(",","",TVMarketSize))) %>%
  rename(team = Team)

# CWS/CHC, LAA/LAD, NYM/NYY, OAK/SFG all share TV Markets so must proportion by TV Rating
clean_tms <- clean_tms %>%
  mutate(proportions = case_when(
    team == "Chicago Cubs" ~ (4.43 / (4.43 + 1.03)),
    team == "Chicago White Sox" ~ (1.03 / (4.43 + 1.03)),
    team == "Los Angeles Angels" ~ (1.43 / (1.43 + 1.96)),
    team == "Los Angeles Dodgers" ~ (1.96 / (1.43 + 1.96)),
    team == "New York Mets" ~ (2.47 / (2.47 + 3.34)),
    team == "New York Yankees" ~ (3.34 / (2.47 + 3.34)),
    team == "Oakland Athletics" ~ (0.88 / (0.88 + 2.68)),
    team == "San Francisco Giants" ~ (0.88 / (0.88 + 2.68)),
    TRUE ~ 1
    )) %>%
  mutate(propTVMarketSize = round(TVMarketSize * proportions)) %>%
  select(team, propTVMarketSize)
```

## Cleaning Team Performance
```{r}
clean_tp <- tp %>%
  mutate(team = name, 
         year = yearID,
         team = (gsub("Anaheim Angels","Los Angeles Angels of Anaheim", team)),
         team = (gsub("Florida Marlins","Miami Marlins", team))) %>%
  select(year, team, W, L, DivWin, WCWin, LgWin, WSWin, attendance) %>%
  arrange(year, team)
```

## Cleaning Attendance and Revenue dataset 
```{r}
clean_ar <- ar %>%
  mutate(attendance = as.numeric(gsub(",","",Attendance)),
         att_per_game = as.numeric(gsub(",","",Attend.G)),
         payroll = as.numeric(str_replace_all(Est..Payroll, "[^[:alnum:]]", ""))) %>%
  rename(year = Year) %>%
  select(year, Tm, attendance, att_per_game, payroll)

clean_ar <- left_join(clean_ar, clean_tp, by = c("year", "attendance")) %>%
  select(year, team, attendance, att_per_game, payroll)
```

## Combining Datasets
```{r}
combined <- left_join(clean_gt, clean_po, by = c("team", "year")) %>%
  mutate(date = yearmonth(date)) %>%
  select(-year) %>%
  as_tsibble(key = team, index = date)

combined
```

# Building Model
```{r}
model_test <- combined %>%
  filter(team =="Chicago Cubs") %>%
  model(
    snaive = SNAIVE(googlehits),
    ets = ETS(googlehits),
    arima = ARIMA(googlehits)
  )

model_test$arima
```