### 2020-07-27 Write_Up.Rmd

---
title: "MLB Write Up"
author: "Jake Esprabens"
date: "7/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Collection

In order to measure the bandwagon level of MLB fans, we gathered data from google trends to depict the popularity of each team.  This data was gathered using the **gtrendsR** package, which automatically pulls the data from Google Trends API.  The units of the Google Trend data are relative to the maximum