### 2020-07-30 Model_Build.Rmd
Added plot of fitted time series

---
title: "Model Building"
author: "Jake Esprabens"
date: "7/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fable)
library(lubridate)
library(tsibble)
library(feasts)
library(mable)
```

## Loading Data
```{r}
gt <- read.csv("MLB_Google_Trends.csv")
ar <- read.csv("MLB_Attendance_and_Revenue.csv")
po <- read.csv("MLB_Preseason_Odds.csv")
tp <- read.csv("MLB_Team_Performance.csv")
tms <- read.csv("MLB_TV_Market_Size.csv")
```

## Cleaning Google Trends Data
```{r}
# cleaning google trends data by replacing <1 with 0 and and changing type of other variables
clean_gt <- gt %>%
  mutate(date = as.character(date),
         year = as.character(year(date)),
         hits = str_replace(hits, "<1","0"),
         hits = as.numeric(hits)) %>%
  rename(team = teams,
         googlehits = hits) %>%
  select(-X) %>%
  arrange(team)
```

## Cleaning Preseason Odds Data
```{r}
# changed the structure of the data using pivot_longer and then changed certain names of columns
clean_po <- po %>%
  pivot_longer(cols = 2:18, names_to = "year") %>%
  mutate(year = substring(year, 2)) %>%
  rename(team = Team,
         odds = value)
```

## Cleaning TV Market Size
```{r}
# take out all commas in observations
clean_tms <- tms %>%
  mutate_all(na_if,"") %>%
  mutate(TVMarketSize = as.numeric(gsub(",","",TVMarketSize))) %>%
  rename(team = Team)

# CWS/CHC, LAA/LAD, NYM/NYY, OAK/SFG all share TV Markets so must proportion by TV Rating to split the Market size
clean_tms <- clean_tms %>%
  mutate(proportions = case_when(
    team == "Chicago Cubs" ~ (4.43 / (4.43 + 1.03)),
    team == "Chicago White Sox" ~ (1.03 / (4.43 + 1.03)),
    team == "Los Angeles Angels" ~ (1.43 / (1.43 + 1.96)),
    team == "Los Angeles Dodgers" ~ (1.96 / (1.43 + 1.96)),
    team == "New York Mets" ~ (2.47 / (2.47 + 3.34)),
    team == "New York Yankees" ~ (3.34 / (2.47 + 3.34)),
    team == "Oakland Athletics" ~ (0.88 / (0.88 + 2.68)),
    team == "San Francisco Giants" ~ (2.68 / (0.88 + 2.68)),
    TRUE ~ 1
    )) %>%
  mutate(propTVMarketSize = round(TVMarketSize * proportions)) %>%
  select(team, propTVMarketSize)
```

## Cleaning Team Performance
```{r}
# changed Angels, Marlins, and Expos to all their updated names / cities
# changed the N and Y to 0's and 1's to indicate whether each team won their division, wild card game, league, and world series
clean_tp <- tp %>%
  mutate(team = name, 
         year = yearID,
         team = (gsub("Anaheim Angels","Los Angeles Angels of Anaheim", team)),
         team = (gsub("Florida Marlins","Miami Marlins", team)),
         team = (gsub("Montreal Expos","Washington Nationals", team))) %>%
  mutate(year = as.character(year),
         team = as.factor(team)) %>%
  mutate(DivWin = as.factor(case_when(DivWin == "N" ~ 0, DivWin == "Y" ~ 1)),
         WCWin = as.factor(case_when(WCWin == "N" ~ 0, WCWin == "Y" ~ 1)),
         LgWin = as.factor(case_when(LgWin == "N" ~ 0, LgWin == "Y" ~ 1)),
         WSWin = as.factor(case_when(WSWin == "N" ~ 0, WSWin == "Y" ~ 1))) %>%
  mutate(winpct = round(W / (W + L), digits = 3)) %>%
  select(year, team, winpct, DivWin, WCWin, LgWin, WSWin, attendance) %>%
  arrange(year, team)
```

## Cleaning Attendance and Revenue dataset 
```{r}
# took out commas in all numeric observations
clean_ar <- ar %>%
  rename(year = Year) %>%
  mutate(attendance = as.numeric(gsub(",","",Attendance)),
         attgame = as.numeric(gsub(",","",Attend.G)),
         payroll = as.numeric(str_replace_all(Est..Payroll, "[^[:alnum:]]", "")),
         year = as.character(year),
         team = as.factor(Tm)) %>%
  select(year, Tm, attendance, attgame, payroll)

# needed to get full names of teams so used team names from previous dataframe
clean_ar <- left_join(clean_ar, clean_tp, by = c("year", "attendance")) %>%
  select(year, team, attendance, attgame, payroll)
```

## Combining Datasets
```{r}
# combining all dataframes together
combined <- left_join(clean_gt, clean_po, by = c("team", "year")) # combine google trends and preseason odds data
combined <- left_join(combined, clean_tms, by = "team") # combine previous dataframe with TVMarketSize data
combined <- left_join(combined, clean_tp, by = c("team", "year")) # combine previous dataframe with team performance data
combined <- left_join(combined, clean_ar, by = c("team", "year")) # combine previous dataframe with attendance and revenue data

# picking out and renaming specific variables
# converting date to a month object to fit in fable / tsibble functions
# convert to tsibble
alldata <- combined %>%
  rename(attendance = attendance.x) %>%
  select(-c(year, attendance.y)) %>%
  mutate(date = yearmonth(date)) %>%
  as_tsibble(key = team, index = date)
```

# Building Model
```{r}
# filtering down to one team
lad <- alldata %>%
  filter(team == "Los Angeles Dodgers") %>%
  arrange(date)

# creating snaive, ets, and arima models
fit <- lad %>%
  model(
    arima = ARIMA(googlehits ~ odds + attgame + winpct) # testing with only arima right now
  )

# ARIMA is used when implementing exogenous regressors

# DIFFERENT FUNCTIONS TO USE ON MABLE
# augment() grabs fitted values and we can plot these
# glance() checks AIC and BIC
# components() allows us to use ggplot to plot aspects of the model / only works on ets and snaive i think

fitted <- augment(fit)

fitted %>%
  ggplot() + 
  geom_line(aes(x = date, y = googlehits), color = "blue") +
  geom_line(aes(x = date, y = .fitted), color = "red")

# small problem is that it can't produce fitted values for any 2020 months because I don't have winpct for those dates / dont know if it matters?

coef <- fit %>%
  select(team, arima) %>%
  coef()

coef
```