### 2020-08-11 KB_modeling.Rmd
still frustrate

---
title: "Hierarchical ARIMA"
author: "Kelly Bodwin"
date: "8/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fable)
library(lubridate)
library(tsibble)
library(feasts)
library(mable)
library(hts)
library(modelr)
```

```{r}
dat <- read_csv(here::here("combined_data.csv"))

dat <- dat %>%
  mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
  mutate(
    playoffs = case_when(
      WSWin ~ "Won WS",
      LgWin ~ "Won League",
      WCWin  ~ "Won Conference",
      DivWin  ~ "Won Division",
      TRUE ~ "Did not qualify"
    ),
    playoffs = forcats::fct_inorder(playoffs),
    phase = case_when(
      month < 04 | month > 10 ~ "Summer",
      month == 04 ~ "Opening",
      month < 10 & month > 04 ~ "Season",
      month == 10 ~ "Playoffs"
    ),
    phase = forcats::fct_reorder(phase, month),
    is_playoffs = as.integer(phase == "Playoffs")
  ) %>%
  mutate_at(vars(month, year, team), funs(factor))

summary(dat$playoffs)
```

**Forever** variables:
- `propTVMarketSize` = # of people in TV market region

**Start of Season** variables:
- `odds` = preseason odds of winning WS
- `payroll` = team's payroll at start of season (?)

**End of Season** variables:
- `attendance` = total attendance
- `attgame` = attendance per game (?)
- `winpct` = win percent of season
- `DivWin`, etc = dummy variables for playoff success

**Monthly** variables:
- Just `googlehits`

Overall mean google hits should be determined by market size.

Annual mean hits should be initially predicted by percieved competitiveness.

Annual mean hits should actually occur based on win percent and playoff success.

Annual mean hits are well measured by attendance.

Monthly hits should be determined by:
- Seasonality
- What happened last month with the team
- Playoff status

Do the "boosts" act as a percent of the possible volume?

---
## Explore data

```{r}
dat %>%
  ggplot() +
  geom_histogram(aes(x = log(googlehits))) +
  facet_wrap(~phase)

dat %>%
  filter(googlehits > 0) %>%
  summarize(mean(log(googlehits+1)))
```

Basically, ghits are very, very skewed, and this is a bit of a problem.

I'm going to do a silly cheat that is actually used a lot, and quantile normalize
this data.

```{r}
quant_norm <- function(vec){

  quants <- rank(vec)/(length(vec) + 1)
  normies <- purrr::map_dbl(quants, qnorm)

  return(normies)

}

dat <- dat %>%
  mutate(
    qn_ghits = quant_norm(googlehits)
  )

```

```{r}
CA_teams <- c("San Francisco Giants", "Los Angeles Dodgers", "San Diego Padres", 
              "Oakland Athletics")

example_teams <- c("Chicago Cubs", "Arizona Diamondbacks", "Los Angeles Dodgers")

## Just makes things a little less extreme.

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = date, y = qn_ghits, color = team)) +
  geom_line()

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = date, y = googlehits)) +
  geom_line()
```

---

```{r}
dat %>%
  group_by(playoffs, month) %>% 
  summarize(googlehits = mean(googlehits)) %>%
  ungroup() %>%
 ggplot(aes(x = month, y = googlehits)) +
  geom_point() +
  facet_wrap(~playoffs)

dat %>%
  group_by(playoffs, month) %>% 
  summarize(googlehits = mean(qn_ghits)) %>%
  ungroup() %>%
 ggplot(aes(x = month, y = googlehits)) +
  geom_point() +
  facet_wrap(~playoffs)

```

---
## Ugly model

```{r}
mod <- lm(qn_ghits ~ propTVMarketSize  + odds + attgame + payroll + is_playoffs*winpct +  is_playoffs*playoffs, data = dat)

summary(mod)

dat %>%
  add_predictions(mod) %>%
  mutate(
    error = qn_ghits - pred
  ) %>%
  filter(team %in% c("Boston Red Sox", "New York Yankees", "Arizona Diamondbacks")) %>%
  #filter(team == "Boston Red Sox") %>%
  ggplot() +
   geom_line(aes(x = date, y = error, color = team))
  #geom_line(aes(x = date, y = qn_ghits), color = "cornflowerblue") +
  #geom_line(aes(x = date, y = pred), color = "deeppink")
```

---
## More exploring

```{r}
by_team <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team) %>%  
  summarize(qn_ghits = mean(qn_ghits),
            propTVMarketSize = max(propTVMarketSize))

by_year <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team, year) %>%  
  summarize(qn_ghits = median(qn_ghits),
            attendance = max(attendance),
            payroll = max(payroll),
            odds = max(odds),
            winpct = max(winpct)) %>%
  ungroup()

by_year %>%
  ggplot(aes(x = payroll, y = qn_ghits)) +
  geom_point()

library(GGally)

by_year %>%
  select(-team, -year) %>%
  ggpairs()
```
---

## Model 1: Overall mean

```{r}
by_team %>%
  filter(team %in% example_teams)

global_means <- lm(qn_ghits ~ propTVMarketSize, data = by_team)

summary(global_means)
```

Takeaway:  More hits in bigger markets, as expected.

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(global_means, var = "ghits_p1") %>%
  mutate(
    ghits_e1 = qn_ghits - ghits_p1
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e1, color = team)) +
  geom_line()
```

... but not much evening out happening here.  Seems that the overall levels probably
have more to do with team success than anything else.

---

## Model 2: Annual means

```{r}
by_year <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team, year) %>%  
  summarize(ghits_e1 = mean(ghits_e1),
            attendance = max(attendance),
            payroll = max(payroll),
            odds = max(odds),
            winpct = max(winpct)) %>%
  ungroup()

annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll, data = by_year) ## this should be an AR model

summary(annual_means)

```

Takeaway:  Payroll and winpct are a good predictor of annual ghits

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(annual_means, var = "ghits_p2") %>%
  mutate(
    ghits_e2 = qn_ghits - ghits_p1 - ghits_p2
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e2, color = team)) +
  geom_line()
```

Much more evened out for Dodgers and Dbacks.  Why not for Cubs?  Probably because
of the later increasingtrend they have.

---

## Model 3: Monthly

```{r}
final_model <- 
  dat  %>%
  lm(ghits_e2 ~ phase + is_playoffs:playoffs, data = .) ## needs to be AR model instead

summary(final_model)
```

Takeaway:  More hits in bigger markets, as expected.

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(final_model, var = "ghits_p3") %>%
  mutate(
    ghits_e3 = ghits_e2 - ghits_p3,
    final_preds = ghits_p1 + ghits_p2 + ghits_p3
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e3, color = team)) +
  geom_line()
```

Error seems a lot more random, which is the goal!

```{r}
dat %>%
  filter(team %in% CA_teams) %>%
  ggplot(aes(x = date, y = ghits_e3, color = team)) +
  geom_line()
```
```{r}
dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = ghits_e3)) +
  geom_histogram() +
  facet_wrap(~team)
```

```{r}
dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = qn_ghits, y = final_preds, color = phase)) + 
  geom_point()

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Arizona Diamondbacks") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Los Angeles Dodgers") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Houston Astros") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")
```

Seems to be underpredicting heavily when teams are bad.  Hmmmm.  Maybe the AR model will help that.

---

# Next: A non-ARIMA hierarchical model

```{r}
lump_mod <- lm(qn_ghits ~ payroll + (odds + winpct)*phase + is_playoffs:playoffs, data = dat)

summary(lump_mod)

dat <- dat %>%
  modelr::add_predictions(lump_mod, var = "lump_preds")

dat %>%
  filter(team == "Los Angeles Dodgers") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = lump_preds), color = "cornflowerblue")
```
This is much better.

```{r}
library(lme4)

lump_mod <- lmer(qn_ghits ~ payroll + odds + winpct + (is_playoffs:playoffs + 1|team), data = dat)

summary(lump_mod)

dat <- dat %>%
  modelr::add_predictions(lmer, var = "lump_preds")

dat %>%
  filter(team == "Los Angeles Dodgers") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = lump_preds), color = "cornflowerblue")
```