### 2020-08-15 KB_modeling.Rmd
Added new dataset of playoff success

---
title: "Hierarchical ARIMA"
author: "Kelly Bodwin"
date: "8/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fable)
library(lubridate)
library(tsibble)
library(feasts)
library(mable)
library(hts)
library(modelr)
```

```{r}
dat <- read_csv(here::here("combined_data.csv"))

## Added new dataset for playoff
playoff_success <- read.csv("MLB_Postseason_Success.csv")

dat <- dat %>%
  # mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
  mutate(
  #   playoffs = case_when(
  #     WSWin ~ "Won WS",
  #     LgWin ~ "Lost WS",
  #     WCWin  ~ "Won Wild Card",
  #     DivWin  ~ "Won Division",
  #     TRUE ~ "No Playoffs"
  #   ),
    # playoffs = forcats::fct_inorder(playoffs),
    phase = case_when(
      month < 04 | month > 10 ~ "Summer",
      month == 04 ~ "Opening",
      month < 10 & month > 04 ~ "Season",
      month == 10 ~ "Playoffs"
    ),
    phase = forcats::fct_reorder(phase, month),
    # is_playoffs = as.integer(phase == "Playoffs")
  ) %>%
  mutate_at(vars(month, year, team), funs(factor))

summary(dat$playoffs)

## playoffs variable
#  Did Not make Playoffs or WC Play-in
#  Played in WC Play-In - Going to need to look up and manually add (Is this only the team that lost the wild card?)
#  Eliminated Round 1 - Going to need to look up and manually add
#  Eliminated Round 2 - Going to need to look up and manually add
#  Lost WS - LgWin but no WSwin
#  Won WS - WSWin 
```

**Forever** variables:
- `propTVMarketSize` = # of people in TV market region

**Start of Season** variables:
- `odds` = preseason odds of winning WS
- `payroll` = team's payroll at start of season (?)

**End of Season** variables:
- `attendance` = total attendance
- `attgame` = attendance per game (?)
- `winpct` = win percent of season
- `DivWin`, etc = dummy variables for playoff success

**Monthly** variables:
- Just `googlehits`

Overall mean google hits should be determined by market size.

Annual mean hits should be initially predicted by percieved competitiveness.

Annual mean hits should actually occur based on win percent and playoff success.

Annual mean hits are well measured by attendance.

Monthly hits should be determined by:
- Seasonality
- What happened last month with the team
- Playoff status

Do the "boosts" act as a percent of the possible volume?

---
## Explore data

```{r}
dat %>%
  ggplot() +
  geom_histogram(aes(x = log(googlehits))) #+
  #facet_wrap(~phase)

dat %>%
  filter(googlehits > 0) %>%
  summarize(mean(log(googlehits+1)))
```

Basically, ghits are very, very skewed, and this is a bit of a problem.

I'm going to do a silly cheat that is actually used a lot, and quantile normalize
this data.

```{r}
quant_norm <- function(vec){

  quants <- rank(vec)/(length(vec) + 1)
  normies <- purrr::map_dbl(quants, qnorm)

  return(normies)

}

dat <- dat %>%
  mutate(
    qn_ghits = quant_norm(googlehits)
  )

```

```{r}
CA_teams <- c("San Francisco Giants", "Los Angeles Dodgers", "San Diego Padres", 
              "Oakland Athletics")

example_teams <- c("Chicago Cubs", "Arizona Diamondbacks", "Los Angeles Dodgers")

## Just makes things a little less extreme.

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = date, y = qn_ghits, color = team)) +
  geom_line()

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = date, y = googlehits)) +
  geom_line()
```

---

```{r}
dat %>%
  group_by(playoffs, month) %>% 
  summarize(googlehits = mean(googlehits)) %>%
  ungroup() %>%
 ggplot(aes(x = month, y = googlehits)) +
  geom_point() +
  facet_wrap(~playoffs)

dat %>%
  group_by(playoffs, month) %>% 
  summarize(googlehits = mean(qn_ghits)) %>%
  ungroup() %>%
 ggplot(aes(x = month, y = googlehits)) +
  geom_point() +
  facet_wrap(~playoffs)

```

---
## Ugly model

```{r}
mod <- lm(qn_ghits ~ propTVMarketSize  + odds + attgame + payroll + is_playoffs*winpct +  is_playoffs*playoffs, data = dat)

summary(mod)

dat %>%
  add_predictions(mod) %>%
  mutate(
    error = qn_ghits - pred
  ) %>%
  filter(team %in% c("Boston Red Sox", "New York Yankees", "Arizona Diamondbacks")) %>%
  #filter(team == "Boston Red Sox") %>%
  ggplot() +
   geom_line(aes(x = date, y = error, color = team))
  #geom_line(aes(x = date, y = qn_ghits), color = "cornflowerblue") +
  #geom_line(aes(x = date, y = pred), color = "deeppink")
```

---
## More exploring

```{r}
by_team <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team) %>%  
  summarize(qn_ghits = mean(qn_ghits),
            propTVMarketSize = max(propTVMarketSize))

by_year <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team, year) %>%  
  summarize(qn_ghits = median(qn_ghits),
            attendance = max(attendance),
            payroll = max(payroll),
            odds = max(odds),
            winpct = max(winpct)) %>%
  ungroup()

by_year %>%
  ggplot(aes(x = payroll, y = qn_ghits)) +
  geom_point()

library(GGally)

by_year %>%
  select(-team, -year) %>%
  ggpairs()
```
---

## Model 1: Overall mean

```{r}
by_team %>%
  filter(team %in% example_teams)

global_means <- lm(qn_ghits ~ propTVMarketSize, data = by_team)

summary(global_means)
```

Takeaway:  More hits in bigger markets, as expected.

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(global_means, var = "ghits_p1") %>%
  mutate(
    ghits_e1 = qn_ghits - ghits_p1
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e1, color = team)) +
  geom_line()
```

... but not much evening out happening here.  Seems that the overall levels probably
have more to do with team success than anything else.

---

## Model 2: Annual means

```{r}
by_year <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team, year) %>%  
  summarize(ghits_e1 = mean(ghits_e1),
            attendance = max(attendance),
            payroll = max(payroll),
            odds = max(odds),
            winpct = max(winpct)) %>%
  ungroup()

annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll, data = by_year) ## this should be an AR model

summary(annual_means)

```

Takeaway:  Payroll and winpct are a good predictor of annual ghits

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(annual_means, var = "ghits_p2") %>%
  mutate(
    ghits_e2 = qn_ghits - ghits_p1 - ghits_p2
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e2, color = team)) +
  geom_line()
```

Much more evened out for Dodgers and Dbacks.  Why not for Cubs?  Probably because
of the later increasingtrend they have.

---

## Model 3: Monthly

```{r}
final_model <- 
  dat  %>%
  lm(ghits_e2 ~ phase + is_playoffs:playoffs, data = .) ## needs to be AR model instead

summary(final_model)
```

Takeaway:  More hits in bigger markets, as expected.

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(final_model, var = "ghits_p3") %>%
  mutate(
    ghits_e3 = ghits_e2 - ghits_p3,
    final_preds = ghits_p1 + ghits_p2 + ghits_p3
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e3, color = team)) +
  geom_line()
```

Error seems a lot more random, which is the goal!

```{r}
dat %>%
  filter(team %in% CA_teams) %>%
  ggplot(aes(x = date, y = ghits_e3, color = team)) +
  geom_line()
```
```{r}
dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = ghits_e3)) +
  geom_histogram() +
  facet_wrap(~team)
```

```{r}
dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = qn_ghits, y = final_preds, color = phase)) + 
  geom_point()

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Arizona Diamondbacks") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Los Angeles Dodgers") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Baltimore Orioles") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")
```

Seems to be underpredicting heavily when teams are bad.  Hmmmm.  Maybe the AR model will help that.

---

# Next: A non-ARIMA hierarchical model

```{r}
lump_mod <- lm(qn_ghits ~ payroll + (odds + winpct)*phase + is_playoffs:playoffs:team, data = dat)

summary(lump_mod)

dat <- dat %>%
  modelr::add_predictions(lump_mod, var = "lump_preds")

dat %>%
  filter(team == "Los Angeles Dodgers") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = lump_preds), color = "cornflowerblue")

dat %>%
  filter(team == "San Diego Padres") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = lump_preds), color = "cornflowerblue")
```
This is much better.