### 2020-08-18 KB_modeling.Rmd
nothing

---
title: "Hierarchical ARIMA"
author: "Kelly Bodwin"
date: "8/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fable)
library(lubridate)
library(tsibble)
library(feasts)
library(mable)
library(hts)
library(modelr)
```

```{r}
dat <- read_csv(here::here("combined_data.csv"))

## Added new dataset for playoff: MLB_Postseason_Success.csv
# broken up into 6 different factors
# - Won WS (1 team)
# - Lost WS (1 team)
# - Lost Semi (2 teams)
# - Lost Quarter (4 teams)
# - Lost WC (2 teams)    NOTE: wildcard game started in 2012 so this variable is only from 2012-2019.
# - No Playoffs (22 teams from 2004-2011 and 20 teams from 2012-2019)

playoff_success <- read.csv("MLB_Postseason_Success.csv", stringsAsFactors = FALSE) %>%
  mutate_all(na_if,"") %>%
  replace(is.na(.), "No Playoffs")

#combining new palyoff dataset with dat
dat <- left_join(dat, playoff_success, by = c("team", "year")) %>%
  select(-c(DivWin, WCWin, LgWin, WSWin)) # deselecting unnecessary variables

dat <- dat %>%
  # mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
  mutate(
  #   playoffs = case_when(
  #     WSWin ~ "Won WS",
  #     LgWin ~ "Lost WS",
  #     WCWin  ~ "Won Wild Card",
  #     DivWin  ~ "Won Division",
  #     TRUE ~ "No Playoffs"
  #   ),
    playoffs = forcats::fct_inorder(playoffs), # not exactly sure what this line of code does? do we still need?
    phase = case_when(
      month < 04 | month > 10 ~ "Summer",
      month == 04 ~ "Opening",
      month < 10 & month > 04 ~ "Season",
      month == 10 ~ "Playoffs"
    ),
    phase = forcats::fct_reorder(phase, month), # not exactly sure what this line of code does? do we still need?
    is_playoffs = as.integer(phase == "Playoffs")
  ) %>%
  mutate_at(vars(month, year, team), funs(factor))

summary(dat$playoffs)
```

**Forever** variables:
- `propTVMarketSize` = # of people in TV market region

**Start of Season** variables:
- `odds` = preseason odds of winning WS
- `payroll` = team's payroll at start of season (?)

**End of Season** variables:
- `attendance` = total attendance
- `attgame` = attendance per game (?)
- `winpct` = win percent of season
- `DivWin`, etc = dummy variables for playoff success

**Monthly** variables:
- Just `googlehits`

Overall mean google hits should be determined by market size.

Annual mean hits should be initially predicted by percieved competitiveness.

Annual mean hits should actually occur based on win percent and playoff success.

Annual mean hits are well measured by attendance.

Monthly hits should be determined by:
- Seasonality
- What happened last month with the team
- Playoff status

Do the "boosts" act as a percent of the possible volume?

---
## Explore data

```{r}
dat %>%
  ggplot() +
  geom_histogram(aes(x = log(googlehits))) #+
  #facet_wrap(~phase)

dat %>%
  filter(googlehits > 0) %>%
  summarize(mean(log(googlehits+1)))
```

Basically, ghits are very, very skewed, and this is a bit of a problem.

I'm going to do a silly cheat that is actually used a lot, and quantile normalize
this data.

```{r}
quant_norm <- function(vec){

  quants <- rank(vec)/(length(vec) + 1)
  normies <- purrr::map_dbl(quants, qnorm)

  return(normies)

}

dat <- dat %>%
  mutate(
    qn_ghits = quant_norm(googlehits)
  )

```

```{r}
CA_teams <- c("San Francisco Giants", "Los Angeles Dodgers", "San Diego Padres", 
              "Oakland Athletics")

example_teams <- c("Chicago Cubs", "Arizona Diamondbacks", "Los Angeles Dodgers")

## Just makes things a little less extreme.

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = date, y = qn_ghits, color = team)) +
  geom_line()

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = date, y = googlehits)) +
  geom_line()
```

---

```{r}
dat %>%
  group_by(playoffs, month) %>% 
  summarize(googlehits = mean(googlehits)) %>%
  ungroup() %>%
 ggplot(aes(x = month, y = googlehits)) +
  geom_point() +
  facet_wrap(~playoffs)

dat %>%
  group_by(playoffs, month) %>% 
  summarize(googlehits = mean(qn_ghits)) %>%
  ungroup() %>%
 ggplot(aes(x = month, y = googlehits)) +
  geom_point() +
  facet_wrap(~playoffs)

```

---
## Ugly model

```{r}
mod <- lm(qn_ghits ~ propTVMarketSize  + odds + attgame + payroll + is_playoffs*winpct +  is_playoffs*playoffs, data = dat)

summary(mod)

dat %>%
  add_predictions(mod) %>%
  mutate(
    error = qn_ghits - pred
  ) %>%
  filter(team %in% c("Boston Red Sox", "New York Yankees", "Arizona Diamondbacks")) %>%
  #filter(team == "Boston Red Sox") %>%
  ggplot() +
   geom_line(aes(x = date, y = error, color = team))
  #geom_line(aes(x = date, y = qn_ghits), color = "cornflowerblue") +
  #geom_line(aes(x = date, y = pred), color = "deeppink")
```

---
## More exploring

```{r}
by_team <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team) %>%  
  summarize(qn_ghits = mean(qn_ghits),
            propTVMarketSize = max(propTVMarketSize))

by_year <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team, year) %>%  
  summarize(qn_ghits = median(qn_ghits),
            attendance = max(attendance),
            payroll = max(payroll),
            odds = max(odds),
            winpct = max(winpct)) %>%
  ungroup()

by_year %>%
  ggplot(aes(x = payroll, y = qn_ghits)) +
  geom_point()

library(GGally)

by_year %>%
  select(-team, -year) %>%
  ggpairs()
```
---

## Model 1: Overall mean

```{r}
by_team %>%
  filter(team %in% example_teams)

global_means <- lm(qn_ghits ~ propTVMarketSize, data = by_team)

summary(global_means)
```

Takeaway:  More hits in bigger markets, as expected.

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(global_means, var = "ghits_p1") %>%
  mutate(
    ghits_e1 = qn_ghits - ghits_p1
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e1, color = team)) +
  geom_line()
```

... but not much evening out happening here.  Seems that the overall levels probably
have more to do with team success than anything else.

---

## Model 2: Annual means

```{r}
by_year <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team, year) %>%  
  summarize(ghits_e1 = mean(ghits_e1),
            attendance = max(attendance),
            payroll = max(payroll),
            odds = max(odds),
            winpct = max(winpct)) %>%
  ungroup()

# NOTE to self: check this again to see if i did it right
# i used the lag function , but examples I see have used dynlm package

annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll + lag(ghits_e1, 1), data = by_year) ## this should be an AR model
# annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll, data = by_year)

summary(annual_means)
```

Takeaway:  Payroll and winpct are a good predictor of annual ghits

UPDATE: What changed after adding the AR term lagging one year?
Looks like odds,  winpct, and the lag are all good predictors of annual ghits now.  Payroll is no longer significant.

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(annual_means, var = "ghits_p2") %>%
  mutate(
    ghits_e2 = qn_ghits - ghits_p1 - ghits_p2
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e2, color = team)) +
  geom_line()
```

Much more evened out for Dodgers and Dbacks.  Why not for Cubs?  Probably because
of the later increasing trend they have.

UPDATE: What changed after adding AR term?
Looks like heavier spikes on the under predictions? Note sure why the spikes are so sharp. Is that okay?  Looks more sporadic because the lag accounts for the yearly cycle.

---

## Model 3: Monthly

```{r}
final_model <- 
  dat  %>%
  lm(ghits_e2 ~ phase + is_playoffs:playoffs + lag(ghits_e2, 1), data = .) ## needs to be AR model instead, monthly
  # lm(ghits_e2 ~ phase + is_playoffs:playoffs, data = .)

summary(final_model)
```

Takeaway:  More hits in bigger markets, as expected.

UDPATE: Why is lost WS NA?  All variables are significant predictors of monthly ghits except winning WS.

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(final_model, var = "ghits_p3") %>%
  mutate(
    ghits_e3 = ghits_e2 - ghits_p3,
    final_preds = ghits_p1 + ghits_p2 + ghits_p3
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e3, color = team)) +
  geom_line()
```

Error seems a lot more random, which is the goal!

UPDATE: What changed when I added the lag on the previous month?
It seems like adding the lag term doesn't change this plot? Might be doing something wrong.

```{r}
dat %>%
  filter(team %in% CA_teams) %>%
  ggplot(aes(x = date, y = ghits_e3, color = team)) +
  geom_line()
```

```{r}
dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = ghits_e3)) +
  geom_histogram() +
  facet_wrap(~team)
```

```{r}
dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = qn_ghits, y = final_preds, color = phase)) + 
  geom_point()

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Arizona Diamondbacks") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Los Angeles Dodgers") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Baltimore Orioles") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")
```

Seems to be underpredicting heavily when teams are bad.  Hmmmm.  Maybe the AR model will help that.

UPDATE: AR did seem to help this underpredicting problem! The California team plot is much more sporadic.  The histograms are centered closer to 0.  And all the predictions seem to be closer than before!

---

# Next: A non-ARIMA hierarchical model

```{r}
lump_mod <- lm(qn_ghits ~ payroll + (odds + winpct)*phase + is_playoffs:playoffs:team, data = dat)

summary(lump_mod)

dat <- dat %>%
  modelr::add_predictions(lump_mod, var = "lump_preds")

dat %>%
  filter(team == "Los Angeles Dodgers") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = lump_preds), color = "cornflowerblue")

dat %>%
  filter(team == "San Diego Padres") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = lump_preds), color = "cornflowerblue")
```

This is much better.